### 2017-08-02 sheets/talk-12/sheet-12.tex

@@

      The rest follows from the correctness of the graphical calculus for general monoidal categories and cancellation of inverses.

@@

    \item
      \[ \begin{tikzcd}
        I \otimes A
          \arrow[rrrrrr, bend left, "\sigma"]
          \arrow[r, "\sigma"]
          \arrow[rrdd, "\rho^{-1}"']
          \arrow[rrrddd, bend right, "\lambda"'] &
        A \otimes I
          \arrow[rrrr, bend left, "\sigma"]
          \arrow[r, "1 \otimes \lambda^{-1}"]
          \arrow[rd, "\rho^{-1}"'] &
        A \otimes (I \otimes I)
          \arrow[rr, "\sigma"]
          \arrow[d, "\alpha^{-1}"] &&
        (I \otimes I) \otimes A
          \arrow[d, "\alpha"'] &
        I \otimes A
          \arrow[l, "\lambda^{-1} \otimes 1"']
          \arrow[ld, "\lambda^{-1}"] &
        A \otimes I
          \arrow[l, "\sigma"']
          \arrow[lldd, "\lambda^{-1}"]
          \arrow[lllddd, bend left, "\rho"] \\
        %
        && (A \otimes I) \otimes I
          \arrow[d, "\sigma \otimes 1"] &&
        I \otimes (I \otimes A)
          \arrow[d, "1 \otimes \sigma"'] \\
        %
        && (I \otimes A) \otimes I
          \arrow[rr, "\alpha"] &&
        I \otimes (A \otimes I) \\
        %
        &&& A
      \end{tikzcd} \]
      The rectangle in the middle represents one hexagon equality.
      The section above it commutes by naturality of $\sigma$.
      The squares to it's right and left commute by naturality of $\lambda$ and $\rho$.
      Everything else commutes trivially or by the coherence theorem for general monoidal categories.
    \end{enumerate}
  \end{enumerate}
\end{answer}

@@

\begin{answer}
  You can define the associator, unitors and symmetry by composition of the corresponding morphisms for products and coproducts as well as distributivity.
  As compositions of natural isomorphisms they will also be natural isomorphisms.
  You can then check the triangle, pentagon and hexagon equalities and show that the symmetry is self-inverse (Application of a proof assistant adviced).

  Another approach is to look at a category that includes $\Set$ as subcategory but makes the monoid structure more obvious.
  We observe that $I$ and $\otimes$ coincide with the terminal object and the product functor from the category of sets with partial functions.
  Liker every terminal object and product, they give rise to a monoidal structure.
  Since the associator, unitors and symmetry are isomorphisms, they are total and contained in $\Set$.
  Thus the monoidal structure exists also $\Set$.
\end{answer}

@@

  We define functions in both directions by
  \[ \begin{tikzpicture}
    \node[phantom morphism] (f_left) at (1, 1) {};
    \node[phantom morphism] (f_right) at (2, 1) {};
    \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
    \node at (f) {$!$};

@@

    \draw[object] (in_L)
      to node[auto] {$L$} (L_mid.south)
      to (L_mid.north)
      to[cap object] (f_left);
    \path (f_left) to node[auto, swap] {$R$} (f_left |- out);
    \draw[object] (in_0) to node[auto] {$0$} (f);
    \draw[object] (f_right) to node[auto, swap] {$0$} (out);
  \end{tikzpicture} \]
  and
  \[ \begin{tikzpicture}
    \node[phantom morphism] (f_left) at (0, 0) {};
    \node[phantom morphism] (f_right) at (1, 0) {};
    \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
    \node at (f) {$!$};

    \draw[object] (0.5, -1) to node[auto] {$0$} (f);
    \draw[object] (f_left) to node[auto] {$L$}  (0, 1);
    \draw[object] (f_right) to node[auto] {$0$}  (1, 1);
  \end{tikzpicture}. \]
  In one direction they are trivially inverse because morphisms to $0$ are unique.
  In the other direction it follows by
  \[
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (f_left) at (1, 1) {};
      \node[phantom morphism] (f_right) at (2, 1) {};
      \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
      \node at (f) {$!$};

      \node[phantom morphism] (g_left) at (1.5, 2) {};
      \node[phantom morphism] (g_right) at (2.5, 2) {};
      \node[morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
      \node at (g) {$!$};

@@

      \draw[object] (in_L)
        to node[auto] {$L$} (L_mid.south)
        to (L_mid.north)
        to[cap object] (f_left);
      \path (f_left) to node[auto, swap] {$R$} (f_left |- g.south);
      \draw[object] (in_0) to node[auto] {$0$} (f);
      \draw[object] (f_right) to node[auto, swap] {$0$} (g);
      \draw[object] (g_left) to node[auto] {$L$} (out_L);
      \draw[object] (g_right) to node[auto] {$0$} (out_0);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (f_left) at (1, 1) {};
      \node[phantom morphism] (f_right) at (2.5, 1) {};
      \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
      \node at (f) {$!$};

      \node[phantom morphism] (g_left) at (2, 2) {};
      \node[phantom morphism] (g_right) at (3, 2) {};
      \node[morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
      \node at (g) {$!$};

@@

      \draw[object] (0, 0)
        to node[auto] {$L$} (L_mid.south)
        to (0, 3);
      \draw[object] (1, 0)
        to node[auto] {$0$} (0_mid.south)
        to (1, 3);
    \end{tikzpicture} } }.
  \]
  In the second step we use again that morphisms to $0$ are unique.
\end{answer}

@@

\begin{exercise}
  Show that in a monoidal category where all idempotents split, if there are morphisms $I \xrightarrow{\eta} R \otimes L$ and $L \otimes R \xrightarrow{\varepsilon} I$ satisfying one of the snake equations, then $L$ has a right dual (not necessarily given by $R$ with $\eta$ and $\varepsilon$).
\end{exercise}

@@

    Assume that we have the snake equality
    \[ \vcenter{ \hbox{ \begin{tikzpicture}
      \draw[object] (0, 2)
        to (0, 1)
        to[cup object] (1, 1)
        to[cap object] (2, 1)
        to (2, 0);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \draw (0, 2) to (0, 0);
    \end{tikzpicture} } }. \]

@@

      \draw[object] (0, -1)
        to (0, 0)
        to[cap object] (1, 0)

@@

        to[cup object] (4, 2)
        to (4, 3);

@@

      \draw[object] (0, -1)

@@

          to[cap object] (3, 1)

@@

        to (3, 0)
        to[cup object] (4, 0)
        to (4, 3);

@@

      \draw[object] (0, -1)

@@

        to (2, 3);

@@

  Hence there are morphisms $g$ and $h$ such that

@@

      \draw[object] (in) to (h) to (g) to (out);

@@

@@

      \draw[object] (in) to (g) to (h) to (out);

@@

      \node[phantom morphism, overlay] (out) at (2, 3) {};

@@

        to (0, 1.5)

@@

        to[cup object] (2, 1.5)

@@

  We use those morphisms to define a unit

@@

      \node[morphism] (g) at (1, 0) {$g$};

@@

        to node[auto, swap] {$R$} (0, 1 |- g.north)
        to (0, 1 |- g.south)
        to[cup object] (g.south)
        (g.north) to node[auto] {$L$} (out_right);

@@

@@

        to node[auto] {$L$} (h)
        to[cap object] (1, 0 |- h.north)
        to (1, 0 |- h.south)
        to node[auto, swap] {$R$} (in_right);

@@

      \node[phantom morphism] (out) at (-1, 2.5) {};

@@

      \node[phantom morphism] (in) at (1, -1.5) {};

@@

      \node[phantom morphism] (out) at (-1, 2.5) {};
      \node[phantom morphism] (h) at (2, 1) {};
      \node[phantom morphism] (g) at (0, 0) {};
      \node[phantom morphism] (in) at (3, -1.5) {};

@@

        to (2, 0.5)

@@

        to[cap object] (0, 0.5)

@@

      \node[phantom morphism] (in) at (2, 0) {};

@@

      \node[phantom morphism, overlay] (in) at (0, -0.5) {};
      \node[morphism] (g) at (2, 3) {$g$};
      \node[phantom morphism, overlay] (out) at (2, 4.5) {};

>>>>

\DeclareMathOperator{\myIn}{in}

@@

      The rest follows from the naturality of $\lambda$ and $\rho$ with the correctness of the graphical calculus for general monoidal categories and cancellation of inverses.

@@

    \item
      \[ \begin{tikzcd}
        I \otimes A
          \arrow[rrrrrr, bend left, "\sigma"]
          \arrow[r, "\sigma"]
          \arrow[rrdd, "\rho^{-1}"']
          \arrow[rrrddd, bend right, "\lambda"'] &
        A \otimes I
          \arrow[rrrr, bend left, "\sigma"]
          \arrow[r, "1 \otimes \lambda^{-1}"]
          \arrow[rd, "\rho^{-1}"'] &
        A \otimes (I \otimes I)
          \arrow[rr, "\sigma"]
          \arrow[d, "\alpha^{-1}"] &&
        (I \otimes I) \otimes A
          \arrow[d, "\alpha"'] &
        I \otimes A
          \arrow[l, "\lambda^{-1} \otimes 1"']
          \arrow[ld, "\lambda^{-1}"] &
        A \otimes I
          \arrow[l, "\sigma"']
          \arrow[lldd, "\lambda^{-1}"]
          \arrow[lllddd, bend left, "\rho"] \\
        %
        && (A \otimes I) \otimes I
          \arrow[d, "\sigma \otimes 1"] &&
        I \otimes (I \otimes A)
          \arrow[d, "1 \otimes \sigma"'] \\
        %
        && (I \otimes A) \otimes I
          \arrow[rr, "\alpha"] &&
        I \otimes (A \otimes I) \\
        %
        &&& A
      \end{tikzcd} \]
      The rectangle in the middle represents one hexagon equality.
      The section above it commutes by naturality of $\sigma$.
      The squares to its right and left commute by naturality of $\lambda$ and $\rho$.
      Everything else commutes trivially or by the coherence theorem for general monoidal categories.
    \end{enumerate}
  \end{enumerate}
\end{answer}

@@

\begin{answer}
  First we show that this describes a monoidal category.
  The tensor product an unit are already given.
  The next step is to define the associator and unitors.
  We ignore some parentheses in our notation and denote by $\myIn_i : A_i \to A_1 + \cdots + A_i + \cdots + A_n$ the injection into the $n$-ary disjoint union for all families of sets $A_1, \ldots, A_n$ and indices $1 \leq i \leq n$.
  We define the associator and unitors by pattern matching:
  \begin{align*}
    \alpha_{A, B, C} &: A + B + (A \times B) + C + \roundddd{\rounddd{A + B + (A \times B)} \times C} \\
      &\qquad \to A + B + C + (B \times C) + \roundddd{A \times \rounddd{B + C + (B \times C)}} \\
    \alpha(\myIn_1(a)) &= \myIn_1(a) \\
    \alpha(\myIn_2(b)) &= \myIn_2(b) \\
    \alpha(\myIn_3(a, b)) &= \myIn_5(a, \myIn_1(b)) \\
    \alpha(\myIn_4(c)) &= \myIn_3(c) \\
    \alpha(\myIn_5(\myIn_1(a), c)) &= \myIn_5(a, \myIn_2(c)) \\
    \alpha(\myIn_5(\myIn_2(b), c)) &= \myIn_4(b, c) \\
    \alpha(\myIn_5(\myIn_3(a, b), c)) &= \myIn_5(a, \myIn_3(b, c)) \\
    \\
    \lambda_{A} &: \emptyset + A + (\emptyset \times A) \to A \\
    \lambda(\myIn_2(a)) &= a \\
    \\
    \rho_{A} &: A + \emptyset + (A \times \emptyset) \to A \\
    \rho(\myIn_1(a)) &= a.
  \end{align*}
  Those are obviously bijections since you can basically swap the terms on both sides of the defining equations and get the definition of the inverse.
  They describe natural transformations because they do not change the elements of $A$, $B$ and $C$ but only repackage them.
  To check the triangle and pentagon equalities, is trivial but tedious because you have to make many case distinctions.
  It is best done with a proof assistant.
  We omit it here.

  We showed that $\emptyset$ and $\otimes$ define a monoidal structure.
  Now we will show that it is even symmetric.
  We define the symmetry as
  \begin{align*}
    \sigma_{A, B} &: A + B + (A \times B) \\
      &\qquad\to B + A + (B \times A) \\
    \sigma(\myIn_1(a)) &= \myIn_2(a) \\
    \sigma(\myIn_2(b)) &= \myIn_1(b) \\
    \sigma(\myIn_3(a, b)) &= \myIn3(b, a).
  \end{align*}
  The symmetry is self-inverse because as above, we can swap the argument pattern with the right side of each defining equation and get the same set of equations back.
  Also as above, the symmetry is natural because it does not modify elements of $A$ or $B$.
  The hexagon equalities are again trivial to show, but the proof needs many case distinctions and is thus omitted here.

  Another approach is to look at a category that includes $\Set$ as subcategory but makes the symmetric monoidal structure more obvious.
  We observe that $I$ and $\otimes$ coincide with the terminal object and the product functor from the category of sets with partial functions.
  Liker every terminal object and product, they give rise to a symmetric monoidal structure.
  Since the associator, unitors and symmetry are isomorphisms, they are total and contained in $\Set$.
  Thus the symmetric monoidal structure exists also in $\Set$.
\end{answer}

@@

  We define functions in both directions by
  \[ \begin{tikzpicture}
    \node[phantom morphism] (f_left) at (1, 1) {};
    \node[phantom morphism] (f_right) at (2, 1) {};
    \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
    \node at (f) {$0_{R \otimes 0}$};

@@

    \draw[object] (in_L)
      to node[auto] {$L$} (L_mid.south)
      to (L_mid.north)
      to[cap object] (f_left);
    \path (f_left) to node[auto, swap] {$R$} (f_left |- out);
    \draw[object] (in_0) to node[auto] {$0$} (f);
    \draw[object] (f_right) to node[auto, swap] {$0$} (out);
  \end{tikzpicture} \]
  and
  \[ \begin{tikzpicture}
    \node[phantom morphism] (f_left) at (0, 0) {};
    \node[phantom morphism] (f_right) at (1, 0) {};
    \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
    \node at (f) {$0_{L \otimes R}$};

    \draw[object] (0.5, -1) to node[auto] {$0$} (f);
    \draw[object] (f_left) to node[auto] {$L$}  (0, 1);
    \draw[object] (f_right) to node[auto] {$0$}  (1, 1);
  \end{tikzpicture}. \]
  Where $0_{R \otimes 0}$ and $0_{L \otimes 0}$ stand for the unique morphisms from the initial object.
  In one direction they are trivially inverse because morphisms from $0$ are unique.
  In the other direction it follows by
  \[
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (f_left) at (1, 1) {};
      \node[phantom morphism] (f_right) at (2, 1) {};
      \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
      \node at (f) {$0_{R \otimes 0}$};

      \node[phantom morphism] (g_left) at (1.5, 2) {};
      \node[phantom morphism] (g_right) at (2.5, 2) {};
      \node[morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
      \node at (g) {$0_{L \otimes 0}$};

@@

      \draw[object] (in_L)
        to node[auto] {$L$} (L_mid.south)
        to (L_mid.north)
        to[cap object] (f_left);
      \path (f_left) to node[auto, swap] {$R$} (f_left |- g.south);
      \draw[object] (in_0) to node[auto] {$0$} (f);
      \draw[object] (f_right) to node[auto, swap] {$0$} (g);
      \draw[object] (g_left) to node[auto] {$L$} (out_L);
      \draw[object] (g_right) to node[auto] {$0$} (out_0);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (f_left) at (1, 1) {};
      \node[phantom morphism] (f_right) at (2.5, 1) {};
      \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
      \node at (f) {$0_{R \otimes 0}$};

      \node[phantom morphism] (g_left) at (2, 2) {};
      \node[phantom morphism] (g_right) at (3, 2) {};
      \node[morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
      \node at (g) {$0_{L \otimes 0}$};

@@

      \draw[object] (0, 0)
        to node[auto] {$L$} (L_mid.south)
        to (0, 3);
      \draw[object] (1, 0)
        to node[auto] {$0$} (0_mid.south)
        to (1, 3);
    \end{tikzpicture} } }.
  \]
  In the second step we use again that morphisms from $0$ are unique.
\end{answer}

@@

\begin{exercise}
  Show that in a monoidal category where all idempotents split, if there are morphisms $I \xrightarrow{\eta} R \otimes L$ and $L \otimes R \xrightarrow{\varepsilon} I$ satisfying the first snake equation, then $L$ has a right dual (not necessarily given by $R$ with $\eta$ and $\varepsilon$).
\end{exercise}

@@

    Assume that we have the snake equality
    \[ \vcenter{ \hbox{ \begin{tikzpicture}
      \draw[object] (2, 2)
        to (2, 1)
        to[cup object] (1, 1)
        to[cap object] (0, 1)
        to (0, 0);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \draw (0, 2) (0, 0);
    \end{tikzpicture} } }. \]

@@

      \draw[object] (4, -1)
        to node[auto, near start] {$R$} (4, 0)
        to[cap object] (3, 0)

@@

        to node[auto, near end] {$R$} (0, 3);

@@

      \draw[object] (4, -1)
        to node[auto, very near start] {$R$} (4, 2)
        to (3, 1)

@@

          to[cap object] (1, 1)

@@

        to[cup object] (0, 0)
        to node[auto, very near end] {$R$} (0, 3);

@@

      \draw[object] (2, -1)
        to node[auto, very near start] {$R$} (2, 2)

@@

        to[cup object] (0, 0)
        to node[auto, very near end] {$R$} (0, 3);

@@

  Since all idempotents split, there are morphisms $R \xrightarrow{g} X$ and $ X\xrightarrow{h} R$ such that

@@

        to node[auto] {$X$} (h)
        to node[auto] {$X$} (out);

@@

      \draw[object] (in) to node[auto, very near start] {$X$} (out);

@@

        to node[auto] {$R$} (g)
        to node[auto] {R$$} (out);

@@

      \node[phantom morphism, overlay] (in) at (2, 0) {};

@@

        to node[auto, near start] {$R$} (2, 1.5)

@@

        to[cup object] (0, 1.5)
        to node[auto, near end] {$R$} (out);

@@

  The object $X$ is right dual to $L$.
  We use the morphisms $g$ and $h$ to define a unit

@@

@@

        to node[auto, swap] {$X$} (g)
        to[cup object] (1, 0 |- g.south)
        to (1, 0 |- g.north)
        to node[auto] {$L$} (out_right);

@@

      \node[morphism] (h) at (1, 1) {$h$};

@@

        to node[auto] {$L$} (0, 0 |- h.south)
        to (0, 0 |- h.north)
        (h.south) to node[auto, swap] {$X$} (in_right);
      % \draw[object] (in_left)
      %   to node[auto] {$L$} (h)
      %   to[cap object] (1, 0 |- h.north)
      %   to (1, 0 |- h.south)
      %   to node[auto, swap] {$R$} (in_right);

@@

      \node[phantom morphism] (out) at (1, 2.5) {};

@@

      \node[phantom morphism] (in) at (-1, -1.5) {};

@@

      \node[phantom morphism] (out) at (3, 2.5) {};
      \node[phantom morphism] (h) at (0, 1) {};
      \node[phantom morphism] (g) at (2, 0) {};
      \node[phantom morphism] (in) at (-1, -1.5) {};

@@

        to (0, 0.5)

@@

        to[cap object] (2, 0.5)

@@

      \node[phantom morphism] (out) at (2, 4) {};

@@

      \node[phantom morphism, overlay] (in) at (2, -0.5) {};
      \node[morphism] (h) at (2, 1) {$h$};
      \node[morphism] (g) at (0, 3) {$g$};
      \node[phantom morphism, overlay] (out) at (0, 4.5) {};
